---
title: "Exercise 3"
output: github_document
---

```{r}
#Importing the libraries
library(arrow)
library(readr)
library(wru)
library(lubridate)
library(ggplot2)
library(tidygraph)
library(igraph)
library(visNetwork)

#Importing the datsets
applications <- read_parquet("C:/Users/33652/Documents/Canada/Mcgill/COURS/Winter semester/Orgaziational Network/app_data_sample.parquet",as_data_frame=TRUE)
edges <- read_csv("C:/Users/33652/Documents/Canada/Mcgill/COURS/Winter semester/Orgaziational Network/edges_sample.csv")

```


Question 1

The first part of this exercise is made with the help of the code given by the prof regarding the race, gender and tenure of the different examiners

```{r setup, include=FALSE}
examiner_names <- applications %>% 
  distinct(examiner_name_first)

examiner_names

examiner_names_gender <- examiner_names %>% 
  do(results = gender(.$examiner_name_first, method = "ssa")) %>% 
  unnest(cols = c(results), keep_empty = TRUE) %>% 
  select(
    examiner_name_first = name,
    gender,
    proportion_female
  )

examiner_names_gender

# remove extra colums from the gender table
examiner_names_gender <- examiner_names_gender %>% 
  select(examiner_name_first, gender)

# joining gender back to the dataset
applications <- applications %>% 
  left_join(examiner_names_gender, by = "examiner_name_first")

# cleaning up
rm(examiner_names)
rm(examiner_names_gender)
gc()


#Finding the race of the examiner

examiner_surnames <- applications %>% 
  select(surname = examiner_name_last) %>% 
  distinct()

examiner_surnames

examiner_race <- predict_race(voter.file = examiner_surnames, surname.only = T) %>% 
  as_tibble()

examiner_race


examiner_race <- examiner_race %>% 
  mutate(max_race_p = pmax(pred.asi, pred.bla, pred.his, pred.oth, pred.whi)) %>% 
  mutate(race = case_when(
    max_race_p == pred.asi ~ "Asian",
    max_race_p == pred.bla ~ "black",
    max_race_p == pred.his ~ "Hispanic",
    max_race_p == pred.oth ~ "other",
    max_race_p == pred.whi ~ "white",
    TRUE ~ NA_character_
  ))

examiner_race

# removing extra columns
examiner_race <- examiner_race %>% 
  select(surname,race)

applications <- applications %>% 
  left_join(examiner_race, by = c("examiner_name_last" = "surname"))

rm(examiner_race)
rm(examiner_surnames)
gc()

#Tenure

examiner_dates <- applications %>% 
  select(examiner_id, filing_date, appl_status_date) 

examiner_dates

examiner_dates <- examiner_dates %>% 
  mutate(start_date = ymd(filing_date), end_date = as_date(dmy_hms(appl_status_date)))

examiner_dates <- examiner_dates %>% 
  group_by(examiner_id) %>% 
  summarise(
    earliest_date = min(start_date, na.rm = TRUE), 
    latest_date = max(end_date, na.rm = TRUE),
    tenure_days = interval(earliest_date, latest_date) %/% days(1)
  ) %>% 
  filter(year(latest_date)<2018)

examiner_dates

applications <- applications %>% 
  left_join(examiner_dates, by = "examiner_id")

rm(examiner_dates)
gc()
```

Question 2

We want to focuse on 2 workgroups. To do so, I check wich 2 workgroups are the ones with the most people in them. 

```{r}
#Let's define the workgroups first
applications$workgroups<-substr(applications$examiner_art_unit,1,3)

#Find 2 workgroups with the most people to compare them 
workgroup_count <- table(applications$workgroups)
sorted_workgroup_count <- sort(workgroup_count, decreasing = TRUE)
head(sorted_workgroup_count, n = 2)
```
It appears that group 162 and 179 have the most people. Those are the 2 workgroups I'm going to focus on for the rest of the assignment. 
An examiner can be mentionned several times in a workgroups depending on the dates or other factors. To avoid disorting the data, I chose to only kept one row for each examiner and to delete duplicates 

```{r}
workgroup_179 <- subset(applications, workgroups == "179")
workgroup_179<- workgroup_179 %>% group_by(examiner_id) %>% slice(1)
workgroup_162 <- subset(applications, workgroups == "162")
workgroup_162<- workgroup_162 %>% group_by(examiner_id) %>% slice(1)

##Let's check the demographic for each group
table(workgroup_179$gender)
```
```{r}
table(workgroup_179$race)
```
```{r}
table(workgroup_162$gender)
```

```{r}
table(workgroup_162$race)
```

Both workgroup are composed by a majority of white person, followed by Asian then Black then Hispanic. Major part of the examiner are also Male. 

I also did some plotting to understang the repartition. 

```{r}
ggplot(workgroup_179,aes(x=race, fill=gender))+
  geom_bar(position="dodge")+
  xlab("Race")+
  ylab("Count")+
  ggtitle("Demographic for workgroup 179")
```
```{r}
ggplot(workgroup_162,aes(x=race, fill=gender))+
  geom_bar(position="dodge")+
  xlab("Race")+
  ylab("Count")+
  ggtitle("Demographic for workgroup 162")
```


Question 3

We now want to look at the centrality scores for people from the 2 workgroups previsouly selected.
First step is to create a new dataset with only the rows from the applications where the examiner id from workgroups 179 and 162. Here again we remove the duplicates.
For the edges as well, we only want to keep the edges with the examiner_id within the new_dataset created. I'm also only keeping the columns I need and renaming the egdes with "from" and "to" to be able to use them later. 

```{r}
#We're only looking at the examiners in our 2 workgroups
new_dataset <- rbind(workgroup_179, workgroup_162)

#Same for the edges dataset we only want to keep the edges going from an examiner id within our subset
edges_df<-edges
edges<edges_df

edges<-edges%>%
  filter(ego_examiner_id %in% new_dataset$examiner_id)%>%
  drop_na()%>%
  mutate(from=ego_examiner_id, to=alter_examiner_id)%>%
  select(from,to)

```

We can then create the nodes dataset

```{r}

#Creating the nodes dataset only with the examiner ids from the selected workgroups
nodes<-select(new_dataset, examiner_id)

```



```{r}
visNetwork(nodes, edges) %>%
  visLegend() %>%
  visEdges(
    arrows = list(
      to = list(
        scaleFactor = 0.75,
        type = "arrow"
      ),
      from = list(
        scaleFactor = 0.75,
        type = "arrow"
      )
    )
  )


```


The nest step is to calculte the centrality scores. I first tried the following code but with no success
```{r}

###Tries to put the degree centrality on the visnetwork reprensetation but there are too many nodes. 
##Calculating the degree centrality 
#g <- graph_from_data_frame(edges, directed = FALSE)

# calculate the degree centrality of each node
#degree <- centr_degree(g)$degree

# add the degree values as a node attribute in your nodes data frame
#nodes$degree_centrality <- degree

# plot the network with degree values as labels
#visNetwork(nodes, edges) %>%
#  visNodes(label = "degree_centrality")
```

I've been able to resolve the problem with the following codes:

Centrality degree 
```{r}
###Centrality degree that works 

graph <- graph_from_data_frame(edges, directed = TRUE)

#degree centrality of each node
deg <- degree(graph)
centrality_table <- data.frame(node = V(graph)$name, degree_centrality = deg)
centrality_table <- centrality_table[order(-centrality_table$degree_centrality),]

# Top 20 nodes with highest centrality
head(centrality_table, n = 20)
```

Closeness degree 
```{r}
#Closeness centrality
closeness<-closeness(graph, mode='out')
closeness_table <- data.frame(node = V(graph)$name, closeness_centrality = closeness)
closeness_table <- closeness_table[order(-closeness_table$closeness_centrality),]

head(closeness_table, n = 20)
```

```{r}
#Betweeness Centrality
betweenness <- betweenness(graph, directed = TRUE)
betweenness_table <- data.frame(node = V(graph)$name, betweenness_centrality = betweenness)
betweenness_table <- betweenness_table[order(-betweenness_table$betweenness_centrality),]
head(betweenness_table, n=20)

```

I then added thos centrality scores to the main dataset to be able to study it better
```{r}

##To better be able to check the data, let's add the centralitys to the examiners dataset
# Create degree centrality data frame
deg_centrality <- data.frame(examiner_id = V(graph)$name, degree_centrality = deg)
close_centrality <- data.frame(examiner_id = V(graph)$name, closeness_centrality = closeness)
between_centrality <- data.frame(examiner_id = V(graph)$name, betweenness_centrality = betweenness)

# Merge new_dataset with degree centrality, closeness centrality, and betweenness centrality data frames
merged_dataset <- merge(new_dataset, deg_centrality, by = "examiner_id")
merged_dataset <- merge(merged_dataset, close_centrality, by = "examiner_id")
merged_dataset <- merge(merged_dataset, between_centrality, by = "examiner_id")
```


I've decided to plot the degree centrality depending on the race depending on the work group.

```{r}
#degree centrality
plot_data <- merged_dataset %>%
  group_by(workgroups, race) %>%
  summarize(degree_centrality = sum(degree_centrality)) %>%
  ungroup()

ggplot(plot_data, aes(x = workgroups, y = degree_centrality, fill = race)) +
  geom_col(position = "dodge") +
  labs(x = "Workgroup", y = "Degree Centrality", fill = "Race") +
  theme_classic()
```

```{r}
#closness centrality 
plot_data <- merged_dataset %>%
  group_by(workgroups, race) %>%
  summarize(closeness_centrality = sum(closeness_centrality)) %>%
  ungroup()

ggplot(plot_data, aes(x = workgroups, y = closeness_centrality, fill = race)) +
  geom_col(position = "dodge") +
  labs(x = "Workgroup", y = "Closeness Centrality", fill = "Race") +
  theme_classic()

```

```{r}
#Betweeness centrality 
plot_data <- merged_dataset %>%
  group_by(workgroups, race) %>%
  summarize(betweenness_centrality = sum(betweenness_centrality)) %>%
  ungroup()

ggplot(plot_data, aes(x = workgroups, y = betweenness_centrality, fill = race)) +
  geom_col(position = "dodge") +
  labs(x = "Workgroup", y = "Betweeness Centrality", fill = "Race") +
  theme_classic()
```

The centrality degree and the betweenness centrality talked to me more than the closeness centrality. One thing we can see is that most of people owning the cenetrality degree and betweeness centrality are white people. This can be explained by the fact that white people represents the major part of thoses groups. What is interesting as well is that it's not what is happening in the closeness centrality. The races showing the highest closeness degree are black and hispanic.It seems like those people can be links between other and even between different groups. They might not have has much connection but hey are as much important. 